---
title: "Covid19 vaccine study Final2020"
author: "Cyann Arevalo"
date: "November 2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid19) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
```{r}
dataF=filter(FinalData,sex=="gal")
```


```{r}
barchartGC(~infected + treatment,data=dataF)
barchartGC(~infected + treatment,data=dataF, type="percent")

```


  We can see here that twice the number of women who got the placebo were infected than the women who received the drug/vaccine. This shows us that the drug is effective.

```{r}
table1 <- xtabs(~infected + treatment, data=dataF)
rowPerc(table1)
colPerc(table1)
```
Here we can also see that among those who were infected, 66.89% had the placebo while only 33.11% had the drug. This again shows that the vaccine was effective in preventing infection.

```{r}
fisher.test(table1)
```
What this result shows is that odds ratio of people who got the vaccine their odds of getting COVID was about 5 to 1, which means the vaccine is effective.

## Analysis for Males

```{r}
dataF=filter(FinalData,sex=="guy")
```
<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

```{r}
barchartGC(~infected + treatment,data=dataF)
barchartGC(~infected + treatment,data=dataF, type="percent")

```

  From this chart, we can see that out of those men who were infected with covid, about 40% had the placebo while almost 60% had the drug. This shows about a 20% difference, so in this case the drug was not effective in preventing covid.

```{r}
table1 <- xtabs(~infected + treatment, data=dataF)
rowPerc(table1)
colPerc(table1)
```
From the numerical summary here, we can see that 41.84% of the males who were infected were given the placebo. 58.16% of the males who were infected were given the drug, this is higher than those males who were infected and given the placebo. Again, based on the percentages, we can see that the drug was not effective in preventing covid.

```{r}
fisher.test(table1)
```
The fisher test result shows is that there was a 1.47 chance (or 47% chance) higher of contracting COVID if the men were given the vaccine than those who were given the placebo. This would indicate that there was an association between the vaccine and contracting the infection.

## Analysis for LGBTQ

```{r}
dataF=filter(FinalData,LGBTQ=="gay")
```

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


```{r}
barchartGC(~infected + treatment,data=dataF)
barchartGC(~infected + treatment,data=dataF, type="percent")

```

  
When looking at this bar chart, we can see that out of the LGBTQ who were infected, more than double of the ones who received the drugs were infected than those who received the placebo. This shows that the drug made them more susceptible to infection.
  

```{r}
table1 <- xtabs(~infected + treatment, data=dataF)
rowPerc(table1)
colPerc(table1)
```
Looking at the numerical summary, we can see that out of the LGBTQ infected, 70.69% of those who took the drug were infected while only 29.39% of those who took the placebo were infected. We see that the drug helped in facilitating the infection.

```{r}
fisher.test(table1)
```
According to the fisher test, the LGBTQ who received the vaccine were at about 2.5 (2.49 to be exact) chances higher of contracting COVID than the LGBTQ who received the placebo. Again bringing us to the conclusion that the vaccine made those who took it more susceptible. 






# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


  
  In the case of women, we can see that women who were given the drug/vaccine were less likely to contract COVID-19 than those who were given the placebo. A total of 33.11% of the women who were infected had been given the vaccine, while 66.89% who were infected were given the placebo. The Fisher's Exact Test showed that the women who were given the vaccine were 50% less likely to get infected. It also showed us a p-value of 2.2e-16, which is less than 0.05, therefore we reject the null hypothesis as the drug/vaccine proved to be effective in preventing COVID-19 for women.
  
  In the case of the men, we can see that out of the men who were infected with COVID-19, 41.84% of the males who were infected were given the placebo while 58.16% of the males who were infected were given the drug. The Fisher's Exact Test for men showed that they had a 1.47 chance higher of contracting COVID if given the vaccine than those who were given the placebo. Fisher's Exact Test showed us as well a p-value of 0.0005801. Although this is less than 0.05 and although we reject the null hypothesis, we conclude that the drug/vaccine was effective in creating the opposite of the desired outcome. This means the drug/vaccine was effective in helping contract COVID-19.
  
  In the case of the LGBTQ, we can see that out of the LGBTQ who took the drug, 70.69% were infected while only 29.39% of those who took the placebo were infected. When we look at the Fisher's Exact Test, we can see that the LGBTQ who were given the drug were at 2.5 chances higher of contracting the infection. Looking at the Fisher's Exact Test we can also see the p-value of 5.741e-06. Similar to the case of the men, the p-value is also less than 0.05 and so we will also reject the null hypothesis here. We can conclude that the vaccine was even more effective in making those who took it susceptible to contract COVID-19.
